---
title: "Lab: Simulation and the Central Limit Theorem"
author: ""
output: rmdformats::readthedown
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE,eval=TRUE)

library(tidyverse)
```


## Introduction

In your coursework, you learned a bit about the **Central Limit Theorem**.

In this lab, you will use **simulation** to illustrate this concept.


## Part One:  Summarize fake data

1. Create a new dataset containing four variables: one that comes from a **Normal**
distribution, one from a **Uniform** distribution, one from a **Binomial** distribution, 
and one from an **exponential** distribution.  

You should **not** use the default options for these distributions; e.g., your
Normal data should not have a mean of 0 or a standard deviation of 1, and your 
Binomial data should not have a probability of 0.5.

Your data should have **30 rows**.


(We did *not* learn about the exponential distribution in lecture; I would like
you to figure out the `r, d, p, q` functions for this new distribution yourself.)

Feel free to make up silly names for your fake variables, and/or to add fake names
or other labels to this dataset, if you are inspired.

```{r}
candy_ratings <- tibble(
  red_vines = rnorm(30, 100, 5),
  twizzlers = runif(30, 60, 70),
  reeses = rbinom(30, 100, .8),
  hersheys = rexp(30, .05)
)
```


2. Calculate the **mean** and **standard deviation** for each of your four variables.

```{r}
candy_ratings %>%
  summarize_all(funs(mean, sd)) 
```


3.  Now repeat steps (1) and (2), using the same distributions,
but instead make **1000** rows in your dataset.

```{r}
candy_ratings <- tibble(
  red_vines = rnorm(1000, 100, 5),
  twizzlers = runif(1000, 60, 70),
  reeses = rbinom(1000, 100, .8),
  hersheys = rexp(1000, .05)
)


candy_ratings %>%
  summarize_all(funs(mean, sd)) 
```

Comment on the means and standard deviations in this section, as compared to 
in (2).

```
They are about the same in both cases, because the data is drawn from the
same distribution, regardless of the number of samples.
```



4.  Make a histogram for each of your four variables, with the underlying 
distribution overlayed on top.

```{r}
candy_ratings %>%
  ggplot(aes(x = red_vines)) +
  geom_histogram(aes(y = ..density..), 
                 col = "grey", fill = "pink",
                 bins = 10) +
  stat_function(fun = dnorm, 
                args = list(100, 5),
                col = "cornflowerblue",
                lwd = 2) +
  theme_classic()


candy_ratings %>%
  ggplot(aes(x = twizzlers)) +
  geom_histogram(aes(y = ..density..), 
                 col = "grey", fill = "pink",
                 bins = 10) +
  stat_function(fun = dunif, 
                args = list(60, 70),
                col = "cornflowerblue",
                lwd = 2) +
  theme_classic()

candy_ratings %>%
  count(reeses) %>%
  mutate(
    pct = n/sum(n),
    dbin = dbinom(reeses, 100, .8)
  ) %>%
  ggplot() +
  geom_col(aes(x = reeses, y = pct), 
                 col = "black", fill = "tan") +
  geom_line(aes(x = reeses, y = dbin),
                col = "cornflowerblue",
                lwd = 2) +
  theme_classic()

candy_ratings %>%
  ggplot(aes(x = hersheys)) +
  geom_histogram(aes(y = ..density..), 
                 col = "black", fill = "tan",
                 bins = 10) +
  stat_function(fun = dexp, 
                args = list(.05),
                col = "cornflowerblue",
                lwd = 2) +
  theme_classic()
```



## Part Two: Generating sample means


1. Write a function called `sample_mean`.  This function should take as input
a vector `vec` and an integer `n`.  It should take a random sample of size `n` from
`vec`, then calculate and return the mean of that subsample.

```{r}
sample_mean <- function(vec, n) {
  
  subsample <- sample(vec, n)
  mean(subsample, na.rm = TRUE)
  
  
}

sample_mean(1:10, 5)
```



2. Write a function called `many_sample_means`.  This function should take as
input a vector `vec`, an integer `n`, and an integer `reps`.  It should perform
the `sample_mean` process many times (`reps`) and return a vector of the results.

```{r}
many_sample_means <- function(vec, n, reps = 1000) {
  
  replicate(reps, sample_mean(vec, n))
  
}

many_sample_means(1:10, 5, 10)
```



3. Write a function called `sample_means_ns`. This function should take as
input a vector `vec` and an integer `reps`, and a vector `ns`.  It should perform
the `many_sample_means` process for each of the values in the `ns` vector.  It should
return a **data frame** with the results.

```{r}
sample_means_ns <- function(vec, ns, reps = 1000) {
  
  my_means <- map(ns, ~many_sample_means(vec, .x, reps))
  
  tibble(
    sample_mean = unlist(my_means),
    n = rep(ns, each = reps)
  )
  
}

sample_means_ns(1:10, 5:7, 10)
```




Include the following in your final R Markdown to show your functions work:

```{r, eval = FALSE}
vec <- runif(1000)

sample_mean(vec, 50)
many_sample_means(vec, reps = 10, n = 50)
sample_means_ns(vec, reps = 10, ns = c(5, 50, 500))

```

## Part Three:  Putting it all together

For any **two** of the four variables in your fake dataset from Part One, do the
following:

1. Use your `many_sample_means` function with `reps = 1000` and `n = 10`.


```{r}
twiz_means <- candy_ratings %>%
  pull(twizzlers) %>%
  many_sample_means(n = 10, reps = 1000)

hersh_means <- candy_ratings %>%
  pull(hersheys) %>%
  many_sample_means(n = 10, reps = 1000)
```


    (a) Make histograms of each of your results (no overlay required)
    
    
```{r}
ggplot(, aes(x = twiz_means)) +
  geom_histogram(col = "grey", fill = "pink") +
  theme_classic()


ggplot(, aes(x = hersh_means)) +
  geom_histogram(col = "black", fill = "tan") +
  theme_classic()
```
    
    
    
    (b) Calculate the mean and standard deviation of each of your results.
    
```{r}
mean(twiz_means)
sd(twiz_means)

mean(hersh_means)
sd(hersh_means)
```
    
    
    

2. Use your `many_sample_means` function with `reps = 1000` and `n = 500`.

```{r}
twiz_means <- candy_ratings %>%
  pull(twizzlers) %>%
  many_sample_means(n = 500, reps = 1000)

hersh_means <- candy_ratings %>%
  pull(hersheys) %>%
  many_sample_means(n = 500, reps = 1000)
```


    (a) Make histograms of each of your results (no overlay required)
    
        
```{r}
ggplot(, aes(x = twiz_means)) +
  geom_histogram(col = "grey", fill = "pink") +
  theme_classic()


ggplot(, aes(x = hersh_means)) +
  geom_histogram(col = "black", fill = "tan") +
  theme_classic()
```
    
    
    (b) Calculate the mean and standard deviation of each of your results.
    
    
```{r}
mean(twiz_means)
sd(twiz_means)

mean(hersh_means)
sd(hersh_means)
```
    
3. Comment on the differences or similarities between (1) and (2)

```
The means of the sample means are the same in both cases.
The standard deviations of the sample means are smaller in (2), by a factor
of about 10.
The histogram of the means from the exponential data (Hershey's) is more symmetric
and bell-curve shaped in (2).
```

4. Use your `sample_means_ns` function to try a variety of values of `n`.  
Calculate the standard deviation of the results for each value of `n`.
Make a plot that shows how the standard deviation of the sample means changes with `n`.

```{r, cache = TRUE}
candy_ratings %>% 
  pull(hersheys) %>%
  sample_means_ns(ns = 5:100, reps = 50) %>%
  group_by(n) %>%
  summarize(std_dev = sd(sample_mean)) %>%
  ggplot(aes(x = n, y = std_dev)) +
  geom_line()
```



## Part Four:  Appreciate the CLT

You have been told that the amount of time you have to wait for a bus from Cal 
Poly to Downtown SLO is `exponential(0.02)`; that is, that the true average
wait time is about 50 minutes.

You think this might be a lie.  In the last 30 days, you have waited for the bus 
for 55 minutes on average.

If the bus system is telling the truth about the `exponential(0.02)` distribution, 
how unlucky were you this month?

1. Simulate 10000 random values from the `exponential(0.02)` distribution.  Use
your `many_sample_means` on these values, with `n = 30` and `reps = 1000`.  How
many times did your sample mean exceed 55?

```{r, cache = TRUE}
rand_vals <- rexp(10000, 0.02)

my_sample_means <- many_sample_means(rand_vals, n = 30, reps = 1000)

sum(my_sample_means > 55)
```


2. Use the Central Limit Theorem to assume that a sample mean of exponentially 
distributed values is Normally distributed, with mean $50$ and standard deviation 
$50/\sqrt{n}$.  Find the probability that a sample mean exceeds 55.

```{r}
1 - pnorm(55, mean = 50, sd = 50/sqrt(30))
```


3. Comment on (1) and (2).  Were the answers similar?  Do you believe that bus wait times
really are distributed `exponential(0.02)`?

```
Yes, the answers were similar!

It looks like the chances of an average monthly wait time of 55 minutes or more 
is about 29% - so it's not too surprising.  It's plausible that the proposed
distribution of wait times is true.
```

