---
title: "Lab: Simulation and the Central Limit Theorem"
author: ""
output: rmdformats::readthedown
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE,eval=TRUE)

library(tidyverse)
```


## Introduction

In your coursework, you learned a bit about the **Central Limit Theorem**.

In this lab, you will use **simulation** to illustrate this concept.


## Part One:  Summarize fake data

1. Create a new dataset containing four variables: one that comes from a **Normal**
distribution, one from a **Uniform** distribution, one from a **Binomial** distribution, 
and one from an **exponential** distribution.  

You should **not** use the default options for these distributions; e.g., your
Normal data should not have a mean of 0 or a standard deviation of 1, and your 
Binomial data should not have a probability of 0.5.

Your data should have **30 rows**.


(We did *not* learn about the exponential distribution in lecture; I would like
you to figure out the `r, d, p, q` functions for this new distribution yourself.)

Feel free to make up silly names for your fake variables, and/or to add fake names
or other labels to this dataset, if you are inspired.

2. Calculate the **mean** and **standard deviation** for each of your four variables.

3.  Now repeat steps (1) and (2), using the same distributions,
but instead make **1000** rows in your dataset.

Comment on the means and standard deviations in this section, as compared to 
in (2).


4.  Make a histogram for each of your four variables, with the underlying 
distribution overlayed on top.


## Part Two: Generating sample means


1. Write a function called `sample_mean`.  This function should take as input
a vector `vec` and an integer `n`.  It should take a random sample of size `n` from
`vec`, then calculate and return the mean of that subsample.


2. Write a function called `many_sample_means`.  This function should take as
input a vector `vec`, an integer `n`, and an integer `reps`.  It should perform
the `sample_mean` process many times (`reps`) and return a vector of the results.


3. Write a function called `sample_means_ns`. This function should take as
input a vector `vec` and an integer `reps`, and a vector `ns`.  It should perform
the `many_sample_means` process for each of the values in the `ns` vector.  It should
return a **data frame** with the results.

For example, if `ns <- c(5, 50, 500)` and `reps = 2`, you would return something
like:

```{r, echo = FALSE}
tibble(
  sample_mean = rnorm(6, 3, 5),
  n = rep(c(5, 50, 500), 2)
)
```


Include the following in your final R Markdown to show your functions work:

```{r, eval = FALSE}
vec <- runif(1000)

sample_mean(vec, 50)
many_sample_means(vec, reps = 10, n = 50)
sample_means_ns(vec, reps = 10, ns = c(5, 50, 500))

```

## Part Three:  Putting it all together

For any **two** of the four variables in your fake dataset from Part One, do the
following:

1. Use your `many_sample_means` function with `reps = 1000` and `n = 10`.  

    (a) Make histograms of each of your results (no overlay required)
    (b) Calculate the mean and standard deviation of each of your results.
    

2. Use your `many_sample_means` function with `reps = 1000` and `n = 500`.  

    (a) Make histograms of each of your results (no overlay required)
    (b) Calculate the mean and standard deviation of each of your results.
    
    
3. Comment on the differences or similarities between (1) and (2)


4. Use your `sample_means_ns` function to try a variety of values of `n`.  
Calculate the standard deviation of the results for each value of `n`.
Make a plot that shows how the standard deviation of the sample means changes with `n`.




## Part Four:  Appreciate the CLT

You have been told that the amount of time you have to wait for a bus from Cal 
Poly to Downtown SLO is `exponential(0.02)`; that is, that the true average
wait time is about 50 minutes.

You think this might be a lie.  In the last 30 days, you have waited for the bus 
for 55 minutes on average.

If the bus system is telling the truth about the `exponential(0.02)` distribution, 
how unlucky were you this month?

1. Simulate 10000 random values from the `exponential(0.02)` distribution.  Use
your `many_sample_means` on these values, with `n = 30` and `reps = 1000`.  How
many times did your sample mean exceed 55?

2. Use the Central Limit Theorem to assume that a sample mean of exponentially 
distributed values is Normally distributed, with mean $50$ and standard deviation 
$50/\sqrt{n}$.  Find the probability that a sample mean exceeds 55.

3. Comment on (1) and (2).  Were the answers similar?  Do you believe that bus wait times
really are distributed `exponential(0.02)`?


## Challenge 1:

The Central Limit Theorem works for the sample **mean**.  Does it work for any
other summary statistics?

Try out at least two other statistics.  Some suggestions:
* the median
* the variance
* the midhinge
* the maximum

Write a very brief argument, using simulation and visualization, about whether or not
the CLT works on the statistic in question.

Upload your writeup **separately**.

## Challenge 2:

Put your 3 functions (`sample_mean`, `many_sample_means`, `sample_means_ns`) 
into an R package in a github repo.  You may copy the `twelvedays` package and
make changes to that infrastructure, to make this easier.