---
title: "Basics: Functions and Troubleshooting"
subtitle: "Week 2"  

output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(flair)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo(
  primary_color = "#26116c",
  secondary_color = "#fd5e53",
  title_slide_text_color = "#fff8e7",
  header_font_google = google_font("Roboto"),
  text_font_google   = google_font("Roboto Condensed"),
  code_font_google   = google_font("Droid Mono")
)
```

class: center, middle

# Functions

---
# Important R Facts

*Fact 1:* **R** is an **open-source** language.

--

*Fact 2:* The power of **R** comes from **packages**.

--

*Fact 3:* **R** is a **functional programming** language.

--

(This means it is designed for using and writing *functions*.)

---
# Anatomy of a function

```{r rnorm1}
rnorm(10, mean = 5, sd =3)
```

---
# Anatomy of a function

```{r rnorm, eval = FALSE}
rnorm(10, mean = 5, sd =3)
```

--
Function name:

```{r, echo = FALSE}
decorate("rnorm", eval = FALSE) %>%
  flair("rnorm")
```

--

Required arguments:


```{r, echo = FALSE}
decorate("rnorm", eval = FALSE) %>%
  flair("10")
```

(we often don't use the names for required arguments, but `n = 10` would be okay too)

---
# Anatomy of a function

Optional arguments:


```{r, echo = FALSE}
decorate("rnorm", eval = FALSE) %>%
  flair("mean = 5") %>%
  flair("sd = 5")
```

(*always* use the argument names for optional arguments)

---
class: center, middle

# Troubleshooting Errors

---

# Syntax Errors

* Did you leave off a parenthesis?

```{r, echo = FALSE}
decorate('rnorm(10, mean = 5', eval = FALSE) %>%
  flair("(")
```

```{r, echo = FALSE, error = TRUE}

rnorm(10, mean = 5

```


---

# Syntax Errors

* Did you leave off a comma?

```{r, echo = FALSE, error = TRUE}
decorate('rnorm(10 mean = 5)', eval = FALSE) %>%
  flair(" mean")
```

```{r, echo = FALSE, error = TRUE}

rnorm(10 mean = 5)

```


---

# Syntax Errors

* Did you make a typo?


```{r, echo = FALSE}
decorate('rnrom(10, mean = 5)', eval = FALSE) %>%
  flair("rnrom")
```

```{r, echo = FALSE, error = TRUE}

rnrom(10, mean = 5)

```

---
# Object type errors

* Are you using the right **input** that the function expects?

```{r, echo = FALSE}
decorate("rnorm('hi', mean = 5)", eval = FALSE) %>%
  flair("'hi'")
```

```{r, echo = FALSE, error = TRUE}

rnorm('hi', mean = 5)

```

---
# Object type errors

* Are you expecting the right **output** of the function?

```{r closure, echo = FALSE, error = TRUE}

my_obj <- rnorm(10, mean = 5)

my_obj(5)

```

```{r, echo = FALSE}
decorate("closure") %>%
  flair("my_obj(5)")
```

---
class: center, middle

# Errors, warnings, and messages

---
# Messages

Just because you see scary red text, this does **not** mean something went wrong!  

This is just **R** communicating with you.

--

For example, you often see:

```{r}
library(tidyverse)
```

---
# Warnings

Often, **R** will give you a **warning**.

--

This means that your code *did* run...

--

...but you probably want to make sure it succeeded.

---
# Warnings


```{r}
my_vec <- c("a", "b", "c")

my_new_vec <- as.integer(my_vec)
```

--

```{r}
my_new_vec
```

---
# Errors

If the word **Error** appears in your message from **R**, then you have a problem.

--

This means your code **could not run**!

--

```{r, error = TRUE}
my_vec <- c("a", "b", "c")

my_new_vec <- my_vec + 1
```
---
class: center, middle

## Parlez-vous ERROR?

Error messages can be confusing!

---
# Translations

**R** says...

> Error: Object 'some_obj' not found.

It *probably* means...

> You haven't run the code to create 'some_obj'  OR you have a typo in the name!

--

```{r, error = TRUE}
some_ojb <- 1:10

mean(some_obj)
```

---
# Translations

**R** says...

> Error: Object of type 'closure' is not subsettable.

It *probably* means...

> Oops, you tried to use square brackets on a function

--

```{r, error = TRUE}
mean[1,2]
```

---
# Translations

**R** says...

> Error: Non-numeric argument to binary operator

It *probably* means...

> You tried to do math on data that isn't numeric.

--

```{r, error = TRUE}
"a" + 2
```

---
# Okay, now what?

When all else fails, **Google** your error message.

--

* Leave out the specifics.

--

* Include the function you are using.


---
# Try it...

*What's wrong here?*

```{r, error = TRUE}
matrix(c("a", "b", "c", "d"), num_row = 2)
```

