---
title: "Basics: Structures and Types"
subtitle: "Week 2"  

output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(flair)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo(
  primary_color = "#26116c",
  secondary_color = "#fd5e53",
  title_slide_text_color = "#fff8e7",
  header_font_google = google_font("Roboto"),
  text_font_google   = google_font("Roboto Condensed"),
  code_font_google   = google_font("Droid Mono")
)
```


class: center, middle, inverse
# Data Types

---
# Data Types

Does your data consist of...

--

* numbers?
    + integers?
    + doubles?
    
--

* words?
    + characters/string?
    + factor?
    
--

* `TRUE` or `FALSE` values (*boolean*)?

--

* something else???
    + Dates and times?
    + Images?
    

---
# Data Types: Numeric

```{r}

a <- 5.5

class(a)

typeof(a)

str(a)

```

---
# Data Types: Numeric
  
```{r}

is.numeric(a)

is.double(a)
```

---
# Data Types: Numeric

*What do you think will happen?*

```{r, eval = FALSE}
a <- 5

is.integer(a)
```

--

```{r, echo = FALSE}
a <- 5

is.integer(a)
```

---

# Data Types: Numeric

Integers are special:


```{r, echo = FALSE}
decorate('a <- 5L') %>%
  flair("L")

```

```{r}
is.integer(a)

typeof(a)
```


--

For this class, it won't be important to make the distinction between integers and doubles.
---
# Data Types: Boolean

For Boolean data , you should think of the data as having values `0` (`FALSE`) or `1` (`TRUE`).

```{r}
is_5 <- a == 5

is_5

class(is_5)

as.integer(is_5)

is_5 + 1
```



---
# Data Types: Character

```{r}
my_word <- "Kelly"

class(my_word)

```

--

```{r}

my_sentence <- "Hi, I'm Kelly!"

class(my_sentence)
```
--
Another word for a **character** type object is a **string**.


---
# Data Types: Factors


```{r, echo = FALSE}
library(palmerpenguins)
```

```{r}
head(penguins)
```


---
# Data Types: Factors

Not just *any* string, but specifically defined **categories**.

```{r}
class(penguins$species)

levels(penguins$species)

str(penguins$species)
```


---
# Data Types: Factors

Warning:  Factors are *weird*

```{r}
as.character(penguins$species)
```
---
# Data Types: Factors

Warning:  Factors are *weird*

```{r}
as.integer(penguins$species)
```

---
class: center, middle
# Object Structures

---

# Object Structures

Is your object a...

* function?

* vector?

* list?

* data frame?

* matrix?

---
# Vectors

A **vector** is a set of values that are all the same **type**

```{r, error  = TRUE}
my_vec <- c("a", "b", "c")

str(my_vec)

my_vec <- c(1, 3, 5)

str(my_vec)

my_vec <- c(1, "a")

str(my_vec)
```

---
# Vectors

A vector with only one **element** is called an **atomic vector**.

```{r, error  = TRUE}

my_vec <- 1

str(my_vec)

my_vec[1]

```

---

# Matrices

A **matrix** is a **two-dimensional** set of values that are all the same type.



```{r, error  = TRUE}
my_mat <- matrix(c("a", "b", "c", "d"), nrow = 2, ncol = 2)

str(my_mat)
```

--
Note: If turning  vector into a matrix, be careful with rows and columns!

```{r}
my_mat
```


---
# Lists

A **list** is the the **most general** object.  Any structure and any data type can be stored in a list!

```{r}
my_list <- list(
  a = "Well hi there",
  b = 1:10,
  c = my_mat
)

my_list
```

---
# Lists

```{r}
str(my_list)

my_list$a
```

---
# Lists: brackets

What is the difference between the following two ways of accessing information in a list?

```{r}
my_list[[2]]

my_list[2]
```

---
# Lists: brackets

What is the difference between the following two ways of accessing information in a list?

```{r}
str(my_list[[2]])
```

--

```{r}

str(my_list[2])
```

---
class: center, middle, inverse

# Check-In: Object structures and data types
