---
title: "Regular Expressions"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = "../../Images/icons") 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(couRsework)
library(flair)
library(lubridate)
```


In this lesson, you will learn to:

* Write regular expressions to match strings

* Use special regular expression characters for general matching.

* Use `stringr()` functions to analyze text with regular expressions

```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "20 min", 
                   reading_time = "0-20 min", 
                   work_time = "90 min", 
                   num_activities = 4)
```


---

## Regular Expressions

```{r, results = "asis", echo = FALSE}
req_vid("Regular Expressions", type = "Required", url = "https://www.youtube.com/embed/zBof1yfwoYE")
```


```{r, results = "asis", echo = FALSE}
req_read("R4DS: String Matching with Regular Expressions", url = "https://r4ds.had.co.nz/strings.html#matching-patterns-with-regular-expressions", type = "Recommended")
```

---

```{r, results = "asis", echo = FALSE}
checkin("A few more symbols")
```

**Question 1:**

Recall that the regular expression `[abc]` matches the characters `a`, `b`, or `c`.  

What does `[^abc]` match?

**Question 2:**

When it is **not** inside square brackets, the `^` symbol means "start of string".

What will be returned by the following?

```{r, eval = FALSE}
my_str <- c("Kelly", "Hi Kelly", "Kelly Bodwin", "Who is Kelly?")
str_subset(my_str, "^Kelly")
```

**Question 3:**

THe `$` symbol in a regular expression means "end of string".

What will be returned by the following?

```{r, eval = FALSE}
my_str <- c("Kelly", "Hi Kelly", "Kelly Bodwin", "Who is Kelly?")
str_subset(my_str, "Kelly$")
```

---

```{r, results = "asis", echo = FALSE}
checkin("Simple Regular Expressions")
```

What will the following outputs be?

```{r, eval = FALSE}
my_str <- "The Dursleys of Number 4 Privet Drive were happy to say that they were perfectly normal, thank you very much."

str_extract_all(my_str, ".*")

str_extract_all(my_str, "\\w")

str_extract_all(my_str, "\\s")

str_extract_all(my_str, "[:alpha:]+")

str_extract_all(my_str, "[:alpha:]*\\.")

str_extract_all(my_str, "[wv]er[ey]")

```

---


```{r, results = "asis", echo = FALSE}
checkin("Complex Regular Expressions")
```


```{r, eval = FALSE}
my_str <- "The Dursleys of Number 4 Privet Drive were happy to say that they were perfectly normal, thank you very much."

str_extract_all(my_str, "[:digit:] ([A-Z][a-z]*)+")

str_extract_all(my_str, "(?<=[:digit:] )[:alpha:]+")

str_extract_all(my_str, "[:digit:].*Drive")

my_str %>%
  str_split() %>%
  str_extract("^[A-Z]")
```

---


```{r, results = "asis", echo = FALSE}
req_vid("Making Word Puzzles", type = "Optional", url = "https://www.youtube.com/watch?v=XldSqU8DbEA")
```

This video is **very** optional.

It uses several functions we won't cover, but it also has some great examples of
simple regular expresions being used with `stringr` functions.

---


```{r, results = "asis", echo = FALSE}
checkin("Text Analysis with Regular Expressions")
```


The file `hamlet_speech.txt`, posted on the course sit, contains the text of a famous speech from the play "Hamlet" by Shakespeare.  Download this file and save it somewhere reasonable.  Read it into R with:

```{r, eval = FALSE}
hamlet <- readLines("hamlet_speech.txt")
```

Answer the following:

* How many words are in the speech?  (Hint: `str_count`)

* How many times does Hamlet reference death or dying?

* How many sentences are in the speech?

* What is the longest word in the speech?

* What is the only capitalized word that does not start a sentence or line?


*Hint:* Right now, your object is a **vector** of type **character**, where each element is a line of the speech.  You may want to use `str_c()` (with appropriate arguments) to turn this into a **single string**.  You may also want to turn it into a vector where each element is **one word**.

Or you may want to do all three!  Different tasks will be easier with different object structures.

