---
title: "Writing R Functions"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = "../../Images/icons") 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(couRsework)
library(flair)
library(lubridate)
```


In this lesson, you will learn to:

* Write your own functions in R

* Make good decisions about function arguments and returns

* Include side effects and/or error messages in your functions

* Use good R code style

```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "30 min", 
                   reading_time = "10-30 min", 
                   work_time = "60-90 min", 
                   num_activities = 4)
```

---

## Writing Functions

---

```{r, results = "asis", echo = FALSE}
req_vid("Writing Functions", type = "Required", url = "https://www.youtube.com/embed/qv_KAY-IBsk")
```


```{r, results = "asis", echo = FALSE}
req_tutorial("Primer: Writing Functions", url = "https://rstudio.cloud/learn/primers/6", type = "Required")
```


```{r, results = "asis", echo = FALSE}
req_read("R4DS: Functions", url = "https://r4ds.had.co.nz/functions.html", type = "Recommended")
```

---

```{r, results = "asis", echo = FALSE}
checkin("Simple Functions")
```

Write a function called `times_seven()` which takes a single argument and multiplies by 7. 

This function should check that the argument is numeric.

This function should also excitedly announce (print) "I love sevens!" if the argument to the function is a 7.

---

```{r, results = "asis", echo = FALSE}
checkin("Predicting Function Behavior")
```

```{r addtwo}
add_or_subtract <- function(first_num, second_num = 2, type = "add") {
  
  if (type == "add") {
    first_num + second_num
  } else if (type == "subtract") {
    first_num - second_num
  } else {
    stop("Please choose `add` or `subtract` as the type.")
  }
  
}
  
```

**Question 1:** What will be returned by each of the following?

i) 1
ii) -1
iii) 30
iv) An error defined in the function `add_or_subtract`
v) An error defined in a different function, that is called from inside `add_or_subtract`

```{r, eval = FALSE}
add_or_subtract(5, 6, type = "subtract")

add_or_subtract("orange")

add_or_subtract(5, 6, type = "multiply")

add_or_subtract("orange", type = "multiply")
```

**Question 2:**

Consider the following code:

```{r}
first_num <- 5
second_num <- 3

result <- 8

result <- add_or_subtract(first_num, second_num = 4)

result_2 <- add_or_subtract(first_num)
```

In your Global Environment, what is the value of...

a) `first_num`
b) `second_num`
c) `result`
d) `result_2`


---

```{r, results = "asis", echo = FALSE}
checkin("Functions and Data Wrangling")
```

Write a function called `find_car_make()` that takes as input the name of a car, and returns only the "make", or the company that created the car.  For example, `find_car_make("Toyota Camry")` should return `"Toyota"` and `find_car_make("Ford Anglica")` should return `"Ford"`.  (You can assume that the first word of a car name is always its make.)

Use your function with the built-in dataset `mtcars`, to create a new column in the data called `make` that gives the company of the car.

*Hint: You should start by using the function `dplyr::rownames_to_column()`, so that the car names are a variable in the data frame instead of row labels.*


---

## Good Code Style

---

Consider the following two chunks of code:

```{r, eval = FALSE}
some_data<-penguins%>%group_by(species)%>%
summarize(av=mean(body_mass_g,0,TRUE))
```

```{r, eval = FALSE}
avg_by_species <-
  penguins %>% 
    group_by(species) %>% 
    summarize(
      avg_mass = mean(body_mass_g, na.rm = TRUE)
    )
```

Which one is easier to read?  The second one, of course!

Part of writing reproducible and shareable code is following good **style guidelines**.  Mostly, this means choosing good object names and using white space in a consistent and clear way.

There is no need to read this full resource, but please skim through it and read some of the examples.


```{r, results = "asis", echo = FALSE}
req_read("Tidyverse style guide", url = "https://style.tidyverse.org/syntax.html", type = "Required")
```

### Summary

Designing functions is somewhat subjective, but there are a few principles that apply:

1. Choose a good, descriptive **names** 
    + Your function name should describe what it does, and usually involves a verb.
    + Your argument names should be simple and/or descriptive.
    + Names of variables in the body of the function should be descriptive.
2. Output should be **very predictable**
    + Your function should always *return* the same object type, no matter what input it gets.
    + Your function should expect certain objects or object types as *input*, 
    and give errors when it does not get them.
    + Your function should give errors or warnings for common mistakes.
    + Default values of arguments should only be used when there is a clear common choice.
3. The body of the function should be **easy to read**.
    + Code should use [good style principles](https://style.tidyverse.org/).
    + There should be occasional comments to explain the purpose of the steps.
    + Complicated steps, or steps that are repeated many times, should be written into **separate functions** (sometimes called *helper functions*).
4. Functions should be **self-contained**.
    + They should not rely on any information besides what is given as input.
    + *(Relying on other _functions_ is fine, though*)
    + They should not alter the Global Environment
    + *(do not put `library()` statements inside functions!)*


---



```{r, results = "asis", echo = FALSE}
checkin("Function Design")
```

Identify five major violations of design principles for the following function:

```{r, eval = FALSE}


ugh <- function(dataset) {
  
  library(tidyverse)
  
  thing <- dataset%>%group_by(Species)%>%summarize(newvar = mean(Sepal.Length))
  
  ggplot(thing)+geom_col(aes(x = Species, y = newvar))+ggtitle("Sepal Lengths")
  
  
}

ugh(iris)
```