---
title: "Hypothesis Testing"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = here::here("Images", "icons")) 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(couRsework)
library(flair)
library(infer)
library(palmerpenguins)
```

In this module, you'll review a few basic hypothesis tests, and learn how to
make R do the calculations for you.

If your memory of hypothesis testing is fresh, you may be able to skip the
review parts of these sections.
You are only expected to know the basic idea behind each test, not every detail.


```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "10-45 min", 
                   reading_time = "30 min", 
                   work_time = "30 min", 
                   num_activities = 2)
```

---

## Review of Hypothesis Testing


```{r, results = "asis", echo = FALSE}
req_vid("Review of Hypothesis Testing", type = "Required", url = "https://www.youtube.com/embed/0oc49DyA3hU")
```

```{r, results = "asis", echo = FALSE}
req_vid("Review of Alternate Hypotheses", type = "Required", url = "https://www.youtube.com/embed/5koKb5B_YWo")
```


```{r, results = "asis", echo = FALSE}
req_vid("Review of p-values", type = "Required", url = "https://www.youtube.com/embed/vemZtEM63GY")
```

---

## t-Tests in R

A **t-test** is used when your hypotheses involve **one or two mean values**, such as
$$ H_0:  \mu_1 = \mu_2 $$
$$ H_a:  \mu_1 > \mu_2 $$

Functions: `t.test()` in base R, or `t_test()` in the `infer` package.

```{r}
t.test(bill_length_mm ~ sex, data = penguins)
```

```{r}
penguins %>%
  t_test(
    response = bill_length_mm,
    explanatory = sex,
    alternative = "two-sided"
  )
```

---

```{r, results = "asis", echo = FALSE}
checkin("t-tests")
```

State (in words) the null and alternate hypotheses for the test in the code above.

State your conclusion.

---

## Chi-Square tests in R

A **Chi-Square** Test is used when your hypotheses involve counts or percents.

Similarly, one option is `chisq.test()` in base R, which needs a *two-way table* 
as input.  

The other option is `chisq_test()` in `infer`, which takes a data frame
and variables as input.  Be careful, though - the variables must be categorical
to be appropriate for a chi squared test.



```{r}
my_table <- mtcars %>%
  janitor::tabyl(cyl, gear)

chisq.test(my_table[,-1])
```
```{r}
mtcars %>%
  mutate(
    cyl = factor(cyl),
    gear = factor(gear)
  ) %>%
  chisq_test(
    response = gear,
    explanatory = cyl
  )
```

```{r, results = "asis", echo = FALSE}
checkin("Chi-Squared Tests")
```

a) Why did we include the `[,-1]` in the first code chunk?

b) Why did we include the `mutate()` step in the second code chunk?

c) What happens if you swap the response and explanatory variable in the second
code chunk?

d) What do you conclude from this test?



---

## Permutation tests

The tests above, and other like them, assume a *distribution* of your test statistic.

We assume that a difference of sample means is approximately Normal (t), because
of the Central Limit Theorem. There is also underlying math involved in showing
that the test statistic for the Chi-Square test has - you guessed it! - a Chi-Square
distribution.

These are called **parametric tests**.

However, sometimes we don't feel comfortable that all our assumptions are met to
assume a distribution, or perhaps we are interested in a test statistic that
does not have an easy-to-derive distribution.  In these cases. we might want to
use a **nonparametric test**.

(Bootstrapping is a form of *nonparametric analysis*!)

The **permutation test** relies on random resampling of the data to determine how
"extreme" the original data is.

```{r, results = "asis", echo = FALSE}
req_vid("Illustration of a Permutation Test", type = "Required", url = "https://www.youtube.com/embed/GmvpsJHGCxQ")
```

(Stop at 8 minutes in.)


```{r, results = "asis", echo = FALSE}
req_read("The infer package", type = "Required", url = "https://teachdatascience.com/infer/")
```

```{r, results = "asis", echo = FALSE}
req_read("ModernDive: Hypothesis Testing", type = "Optional", url = "https://bookdown.org/cteplovs/ismaykim/10-hypothesis-testing.html")
```



---

```{r, results = "asis", echo = FALSE}
checkin("Permutation Tests")
```

```{r permute, include = FALSE, warning = FALSE}
 my_test <- penguins %>% 
  specify(response = bill_length_mm, explanatory = sex) %>%
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("male", "female"))


visualize(my_test) 
```

```{r, echo = FALSE}
decorate("permute") %>%
  mask("specify") %>%
  mask("hypothesize") %>%
  mask("generate") %>%
  mask("calculate") %>%
  mask("visualize") %>%
  mask("independence", background = "pink") %>%
  mask("permute", background = "pink") %>%
  mask("diff in means", background = "pink")
```


a) Fill in the yellow blanks with the appropriate functions from the `infer` package.

b) Fill in the pink blanks with the appropriate arguments.