---
title: "Linear Regression"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = here::here("Images", "icons")) 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(couRsework)
library(flair)
library(infer)
library(palmerpenguins)
```

In this module, you'll learn to compute linear regression models in R.  Feel free
to skip review sections if you are confident in your knowledge.


```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "20 min", 
                   reading_time = "0 min", 
                   work_time = "30 min", 
                   num_activities = 3)
```

---

## Review of Linear Regression


```{r, results = "asis", echo = FALSE}
req_vid("Review of Linear Regression I", type = "Required", url = "https://www.youtube.com/embed/PaFPbb66DxQ")
```

```{r, results = "asis", echo = FALSE}
req_vid("Review of Linear Regression II", type = "Recommended", url = "https://www.youtube.com/embed/nk2CQITm_eo")
```


---

## Linear Regression in R

```{r}
library(palmerpenguins)

penguins %>%
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  geom_point() +
  stat_smooth(method = "lm")


my_model <- penguins %>%
  lm(bill_length_mm ~ bill_depth_mm, data = .)

my_model

summary(my_model)

broom::tidy(my_model)
```

---

```{r, results = "asis", echo = FALSE}
checkin("Linear Regression")
```

**Question 1: Code**

a) What is the `data = .` argument in the `lm()` function?

b) What happens if you switch the order of `bill_length_mm` and `bill_depth_mm`
in the `lm()` formula?

c) What object type was returned by `summary()`?  What about by `tidy()`?


**Question 2: Interpreation**

a) What is the equation for the regression line?

b) Penguin Bob has a bill that is 5mm deeper than Penguin Judy.  How much longer
do you expect Penguin Bob's bill to be?

c) Is the relationship between bill length and bill depth statistically significant?


**Question 3: A more complex model**

Run the following code, and explore the results:

```{r}
my_model_2 <- penguins %>%
  lm(bill_length_mm ~ bill_depth_mm:species, data = .)

my_model_3 <- penguins %>%
  lm(bill_length_mm ~ bill_depth_mm*species, data = .)
```


a) Make a plot illustrating `my_model_2`.  *(Hint: what needs to change in the 
aesthetic of the plot above?)*

b) Which model of the three explains the most variance in the response variable?

c) Do the three species of penguin have the same average bill length?  How do you know?

d) Do the three species of penguin have the same bill shape (i.e., the relationship
between length and depth)?  How do you know?