---
title: "Lab: The 12 Days of Christmas"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)

library(tidyverse)
```

## Introduction

The song "12 Days of Christmas", written around 1780, tells the tale of many gifts 
a person receives in the days leading up to Christmas. ([https://en.wikipedia.org/wiki/The_Twelve_Days_of_Christmas_(song)](https://en.wikipedia.org/wiki/The_Twelve_Days_of_Christmas_(song))) 

These gifts repeat and compound; on the first day, the narrator recieves

```
A partridge in a pear tree.
```

On the last day, he or she receives

```
Twelve Drummers Drumming
Eleven Pipers Piping
Ten Lords a Leaping
Nine Ladies Waiting
Eight Maids a Milking
Seven Swans a Swimming
Six Geese a Laying
Five Golden Rings
Four Calling Birds
Three French Hens
Two Turtle Doves
And a Partridge in a Pear Tree
```

Your task will be to help write a package that automatically sings this very
repetitive song.

---

### A Test Dataset

Run the code below to load up a dataset called `xmas` that contains the crucial 
information about the gifts in the song. We will use this dataset to test out
our functions as we work on them.

```{r}
xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")
```

Note that your functions can - and should! - reference each other.  
That is, don't duplicate code; use earlier, smaller functions inside your larger 
functions.

---

## Step One:  Plurals

Note that the gifts are listed in singular: for example, on day 5 the narrator 
receives "5 golden rings", but the entry in the dataset for the gift on day 5 
simply says "ring".  

In your `twelvedays` package, navigate to the file `pluralize_gift.R` in the `R` 
folder. You will find some incomplete code for a function called `pluralize_gift` 
that takes a gift and returns the appropriate plural. 

Complete this function code.

```{r}
pluralize_gift <- function(item){
  
  if (str_detect(item, "oo")) {
    item <- str_replace(item, "oo", "ee")
  } else if (str_detect(item, "y")) {
    item <- str_replace(item, "y", "ies")
  } else {
    item <- paste0(item, "s")
  }
  
}
```


Try your function out on the gifts in the dataset.  
(Consider: is your function *vectorized*?  
It does not have to be, but it may make things simpler!)

```{r, error = TRUE}

pluralize_gift(xmas$Gift.Item)

purrr::map_chr(xmas$Gift.Item, pluralize_gift)

```
```
This function is NOT vectorized.
```

---



## Step Two: Creating sentences

Write a function called `make_phrase` that takes as input the necessary information, 
and returns a phrase. 

```{r}
make_phrase <- function(num, num_word, item, verb, adjective, location) {
  
  ## Replace NAs with blank strings
  
  item <- replace_na(item, "")
  verb <- replace_na(verb, "")
  adjective <- replace_na(adjective, "")
  location <- replace_na(location, "")
  
  
  ## Pluralize if needed, or change num word to "a/an"
  
  vowel_start <- str_sub(item, 1, 1) %>% str_detect("[aeiou]")
  
  if (num > 1) {
    
    item <- pluralize_gift(item)
    
  } else if (vowel_start) {
    
    num_word <- "an"
    
  } else {
    
    num_word <- "a"
    
  }
  
  glue::glue("{num_word} {adjective} {item} {verb} {location}") %>%
    str_squish()
  
  
}
```


```{r}
make_phrase(num = 10, 
            num_word = "ten", 
            item = "lord", 
            verb = "a-leaping", 
            adjective = "", 
            location = "")
```


Try your function out on the `xmas` data, by making a new variable containing the daily phrases.

```{r}
xmas <- xmas %>%
  mutate(
    Num.Word = as.character(english::english(Day)),
    Full.Phrase = pmap_chr(list(Day, Num.Word, Gift.Item, Verb, Adjective, Location), make_phrase)
  )
```

---


## Step Three:  Iteration.

Write a function called `sing_line` that takes as input:

* A dataset

* A number indicating which day to sing about

* The name of a column in the dataset that contains the phrases for each day.

```{r}
sing_line <- function(data, num, phrase_col) {
  
  
  ## Set up intro line
  
  num_word <- english::ordinal(num)
  
  intro <- glue::glue("On the {num_word} day of Christmas, my true love gave to me:")
  
  ## Sing gift phrases
  
  phrases <- data %>%
    pull({{phrase_col}})

  
  phrases[1] <- paste0("and ", phrases[1], ".")
  
  gift_lines <- str_c(phrases[num:1], collapse = ", \n")
  
  ## put it together
  
  glue::glue("{intro} \n{gift_lines}")
  
  
}
```



```{r}
sing_line(xmas, 2, Full.Phrase)

```

#### Try the package

Use the following code to load a dataset that is different from the original
`xmas` song. 

```{r, echo = TRUE}
xmas2 <- read.csv("https://www.dropbox.com/s/ap2hqssese1ki4j/xmas_2.csv?dl=1")

```

```{r}
xmas2 <- xmas2 %>%
  mutate(
    Num.Word = as.character(english::english(Day)),
    Full.Phrase = pmap_chr(list(Day, Num.Word, Gift.Item, Verb, Adjective, Location), make_phrase)
  ) 

map_chr(1:12, ~sing_line(xmas2, .x, Full.Phrase)) %>%
  str_c(collapse = "\n") %>%
  cat()
  
```

