---
title: "Reproducible Deliverables"
subtitle: "Week 2"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo(
  primary_color = "#26116c",
  secondary_color = "#fd5e53",
  title_slide_text_color = "#fff8e7",
  header_font_google = google_font("Roboto"),
  text_font_google   = google_font("Roboto Condensed"),
  code_font_google   = google_font("Droid Mono")
)
```

## Principles of Reproducibile Documents

1. Document is easy to find and organize.

2. References to data and code work for everyone.

3. Package dependency is clear.

4. Code will run the same every time, even if data values change.

5. Analysis process is well-explained and easy to read.

```{r, include = FALSE}
library(tidyverse)
```

---
class: center, middle, invert
# Deliverables in R Markdown

---
# Setup Steps

- In the frontmatter, choose a good title.

- Save your `.Rmd` file, and name it something meaningful (match the title!)

--

- Your first code chunk looks like this:

````
```{r setup, include=FALSE}`r ''`
knitr::opts_chunk$set(echo = TRUE)
```
````

--

- Your second code chunk looks like this:

````
```{r libraries}`r ''`
library(tidyverse)
```
````

---

## Good organization

- Use `#` to make section headings etc.

- Each code chunk is a unified thought.

- Text before and after each code chunk to motivate your code and explain the output.

- Make deliberate decisions about what to display:
    + Does the reader need to see data loading step?
    + Does the reader need to see your code?
    + Does the reader need to see the output?
    
---  

---

## Naming code chunks

- Notice that code chunks can have labels:

````
```{r data}`r ''`
my_data <- read_csv("penguins.csv")
```
````

--

- Lets you jump between chunks

--

- Careful with copy-pasting, duplicate names are not allowed.


---
## Setting chunk options

- `echo`: Should this code be **shown** (echoed)?

- `eval`: Should this code be **run** (evaluated)?

- `include`: Should this code be neither shown nor run?

- `error`: If there is an error, should that display rather than breaking the knit?

- `warning`: If there is a warning, should that display?

- `message`: If there is a message, should that display?

---
## Setting options for ALL chunks

Recall that first setup chunk.

What do you think this does?

````
```{r global_options, eval=FALSE}`r ''`
knitr::opts_chunk$set(fig.width=6, fig.height=4,
                      echo=TRUE, warning=FALSE, 
                      message=FALSE, eval=TRUE)
```
````

--

- (you can always override these for individual chunks)
