---
title: "Directories, Data, and Deliverables (oh my!)"
subtitle: "Week 2"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo(
  primary_color = "#26116c",
  secondary_color = "#fd5e53",
  title_slide_text_color = "#fff8e7",
  header_font_google = google_font("Roboto"),
  text_font_google   = google_font("Roboto Condensed"),
  code_font_google   = google_font("Droid Mono")
)
```

## Principles of Reproducibility

The idea:  You should be able to send your project to someone else, and they can jump in and start working right away.

1. Files are organized and well-named.

2. References to data and code work for everyone.

3. Package dependency is clear.

4. Code will run the same every time, even if data values change.

5. Analysis process is well-explained and easy to read.

```{r, include = FALSE}
library(tidyverse)
```

---
# Dealing with Directories

## What is a directory?

- A **directory** is just a fancy name for a folder.

- Your **working directory** is the folder that R "thinks" it lives in at the moment.

```{r getwd}
getwd()
```

- If you save things you have created, they save to your working directory by default.

---
class: center, middle

# Manage your Class Directory

--

Is it in a place you can easily find it?

--

Does it have an informative name?

--

Are the files inside it well-organized?

---

## Paths

- A **path** describes where a certain file or directory lives.
```{r getwd2}
getwd()
```
This file lives on my computer `C:/`, 

on my account `Users/kbodwin/` 

in my Dropbox `Dropbox/` 

in a series of organized folders

---
class: center, middle

# Try it!

Find the **path** to your class directory.


---

## The beauty of R Projects

An **R Project** is basically a "flag" planted in a certain directory.

When you double click a **.Rproj** file, it does two things:

1. Open RStudio

2. Set the working directory to be wherever the .Rproj file lives.

3. Link to GitHub, if setup (more on that later!)

--
This is great for reproducibility!  

Anyone can copy your directory, and the .Rproj will open properly on their computer.

---
class: center, middle

# Set up your RProj

In RStudio, choose `File > New Project`.

(Save any files you need to.)

Choose **existing directory** and find your class folder.

--

Close **all** of your RStudio windows.

Go to your class folder.

Double click the **.Rproj** file (it looks like a cube!)

--

**You should ALWAYS access your work this way from now on!**

---

## Good practice

- Organize your folders carefully, and name them meaningfully:  
    + `/User/kbodwin/Dropbox/Teaching/` rather than `Desktop/stuff/`
    
- Use **R Projects** liberally - put one in the "main" folder for each project

    + I have a project for every class I teach.
    + I have a project for every data analysis I do.
    + I **never** open RStudio, except by clicking on a project.

---
# Bad practice

If you put something like this at the top of your .Rmd file, I will set your computer on fire:

```{r setwd, eval = FALSE}
setwd("/User/reginageorge/Desktop/R_Class/Lab_1/")
```

- Setting working directory by hand = BAD!  

--

- That directory is specific to you!

--

- (R Markdown files won't knit this code.)

---
class: middle, inverse

# Do this right now:

## 1. Create a R Project in your class directory.

## 2. Add a folder to the directory called "Data"

## 3. Download the *ages* data from the link provided, and put it in your Data folder.


---
class: center, middle

# Location of Data


---
# read.csv()

- The most common way to read data is with `read.csv()`.

- You may supply either a **url** to the dataset, or a **path**.

```{r read.csv, eval = FALSE}
hiphop <- read.csv("https://www.dropbox.com/s/1qqyshx5ikt9zoc/hiphop.csv?dl=1")
## This will work for everyone!

```

--

```{r read.csv2, eval = FALSE}

hiphop <- read.csv("/Users/kbodwin/Dropbox/Teaching/STAT-331-Online/Data/hiphop.csv")
## This will work only on my computer

```

--

```{r read.csv3, eval = FALSE}

hiphop <- read.csv("hiphop.csv")
## This will work if the file is in the same directory as the code (i.e., the project directory)
```

--

```{r read.csv4, eval = FALSE}

hiphop <- read.csv("./Data/hiphop.csv")
## This will work if a folder called "Data" is in the same directory as the code.
```
---
# The **here** package

A great solution to consistency in paths is the `here` package:

```{r, eval = FALSE}
install.packages("here")
```

--

This package thinks that *"here"* means *"the directory of the closest project"*.

```{r}
here::dr_here()
```
---
# The **here** package

Recall that these slides live in a subfolder for Week 2 and for this Lesson:

```{r}
getwd()
```

--
If I wanted to access a dataset from my main Data folder, I'd have to use a relative file path:

```{r, eval = FALSE}
hiphop <- read.csv("../../Data/hiphop.csv")
```
--

If I ever reshuffle my folder structure, I'm in trouble!

---
# The **here** package

Instead of trying to reference data relative to **this** file, let's refference it relative to the **Project** location, which is unlikely to change.


```{r}
here::here("Data", "hiphop.csv")
```

--


```{r, eval = FALSE}
hiphop <- read.csv(here::here("Data", "hiphop.csv"))
```

This will **adapt** to each person's path!

---
class: center, middle

# Reading in data

---

# Common types of data files

- `.csv` :  "Comma-separated"
```
Name, Age
Bob, 49
Joe, 40
```
- `.txt`:  Plain text
    + Could be just text
    + Could be comma-separated data
    + Could be tab-separated, bar-separated, etc.
    + Need to let R know what to look for

- `.xls, .xlsx`:  Microsoft Excel Spreadsheet
    + Common approach: save as `.csv`
    + Nicer approach:  `readxl` package


---
# read.csv() and read.table()

- A more general version of `read.csv()` is `read.table()`.

- Tell R what symbol separates data observations with `sep = `

```{r read.table, eval = FALSE}
hiphop <- read.table("hiphop.csv", sep = ",")
hiphop <- read.table("hiphop.csv", sep = "\t")  
## if tabs between data
```

---
# readr

The tidyverse has some cleaned-up versions in the `readr` and `readxl` packages: 

- `read_csv` works like `read.csv`, with some extra stuff

- `read_tsv` is for tab-separated data

- `read_table` is for any data with "columns" (white space separating)

- `read_delim` is for special "delimiters" separating data

- `read_excel` is specifically for dealing with Excel files

---
# Exercises

The folder "Age_Datasets" on Canvas several datasets containing the names and ages of five men.  The datasets are stored as different file types.  Save these in a reasonable place, then do the following:

1. Read in the dataset `ages.csv`.
2. Read in the dataset `ages_tab.txt`
3. Read in the dataset `ages_mystery.txt`
4. Read in the dataset `ages.xlsx`
5. Both `read.csv` and `read.table` have an optional argument called `stringsAsFactors`.  Figure out what this does.
6. Find a way to use `read_csv` to read `ages.csv` with the variable "Name" as a factor and "Age" as a character.

---
class: center, middle, invert
# Deliverables in R Markdown

---
# General principles

- Give your file a good title, author, and date

- Name your file something meaningful (match the title!)

- Use `#` to make section headings etc.

- Each code chunk is a unified thought.

- Text before and after each code chunk to motivate your code and explain the output.

- Make deliberate decisions about what to display:
    + Does the reader need to see data loading step?
    + Does the reader need to see your code?
    + Does the reader need to see the output?


---    
# Naming code chunks

- Lets you jump between chunks

- Careful with copy-pasting, duplicate names are not allowed.

(demo)

---
# Setting chunk options

- `echo`: Should this code be **shown** (echoed)?

- `eval`: Should this code be **run** (evaluated)?

- `include`: Should this code be neither shown nor run?

- `error`: If there is an error, should that display rather than breaking the knit?

- `warning`: If there is a warning, should that display?

- `message`: If there is a message, should that display?

---
## Setting options for ALL chunks

What does this do?

```{r global_options, eval=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4,
                      echo=TRUE, warning=FALSE, 
                      message=FALSE, eval=TRUE)
```


- (you can always override these for individual chunks)
