---
title: "Packages"
subtitle: "Week 2"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo(
  primary_color = "#26116c",
  secondary_color = "#fd5e53",
  title_slide_text_color = "#fff8e7",
  header_font_google = google_font("Roboto"),
  text_font_google   = google_font("Roboto Condensed"),
  code_font_google   = google_font("Droid Mono")
)
```

class: center, middle

# Installing packages

---

## install.packages()

To use a package, it must first be *installed* to your copy of R.

This only needs to happen **ONCE** per computer or server.

```{r install packages, eval = FALSE}
install.packages("tidyverse")
```

---

## update.packages()

If a computer you are using has the package, but you aren't sure if it's up to date:

```{r update packages, eval = FALSE}
update.packages("tidyverse")
```

to install the most recent version.

---
class: center, middle, inverse

# NEVER EVER EVER EVER PUT "install.packages" or "update.packages" INTO YOUR R MARKDOWN FILE!!!!!

---

class: center, middle

# Loading Packages

---

# library() and require()

- A package must only be *installed* once.

- But it must be *loaded* every time, to tell R it is needed for this analysis.

- `library` gives an **error** if the package isn't installed

- `require` gives a `TRUE/FALSE` value depending on if the package was available

- For now, you should always use `library`

```{r library, eval = FALSE}
library(tidyverse)

# notice:  no quotes around package name
```

---
class: center, middle, inverse

# Three annoying complications of packages

---
class: center, middle

# Versions

---

## Versions

- From time to time, packages get updated

- This is good!

- But, it can mess with your code, which depended on the older version.

- Outside the scope of this class; just be aware.


---
class: center, middle

# Dependencies

---

## Dependencies

- A *dependency* is when one package makes use of another.

- For example, if you create a package that uses the function `group_by()`, your package **depends** on `dplyr`

- Typically these get loaded automatically!

- Occasionally, something goes weird and you need to manually install the dependency.


---
class: center, middle

# Conflicts

---


## Conflicts

- Sometimes, packages choose the same name for their functions.  This is called a *conflict*.

- The version loaded first takes precedence.  For example...

```{r conflicts1, message = FALSE}
dat <- data.frame(bob = 1:10, joe = 11:20)

library(MASS)
library(dplyr)
```

```{r, error = TRUE}
dat %>% select(bob)
```

---
## Conflicts

```{r, echo = FALSE, message = FALSE, warning = FALSE}
detach("package:dplyr", unload=TRUE)
detach("package:MASS", unload=TRUE)
```

```{r conflicts2, message = FALSE}
dat <- data.frame(bob = 1:10, joe = 11:20)

library(dplyr)
library(MASS)
```

```{r, error = TRUE}
dat %>% select(bob)
```

---

## Conflicts

But this is super annoying, because...

```{r, echo = FALSE, message = FALSE, warning = FALSE}
detach("package:dplyr", unload=TRUE)
detach("package:MASS", unload=TRUE)
```

```{r conflicts3, eval = FALSE}
dat <- data.frame(bob = 1:10, joe = 11:20)

library(dplyr)  # dplyr loaded here
library(MASS)
library(dplyr)  # does nothing, because it sees we already have dplyr
```

```{r, error = TRUE}
dat %>% select(bob)
```

---

## Conflicts

ARG!

Best solution:  Be explicit about your intended package using `::`.

```{r conflicts4, message = FALSE, eval = FALSE}
dat <- data.frame(bob = 1:10, joe = 11:20)

library(dplyr) 
library(MASS)

dat %>% dplyr::select(bob)
```

---
## Conflicts

Another solution: The `conflicted` package

```{r}
library(conflicted)

conflict_prefer("select", "dplyr")
```

