---
title: "Reshaping and Joining Data Frames"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = "../../Images/icons") 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(couRsework)
library(flair)
library(lubridate)
```


In this lesson, you will learn to:

* **Reshape** your dataset by pivoting *longer* or *wider*

* Use **mutating joins** and **filtering joins** to combine different datasets with related information.

* **Decompose** a complicated problem into small, manageable steps.

```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "20 min", 
                   reading_time = "20-60 min", 
                   work_time = "60 min", 
                   num_activities = 3)
```


---

## Reshaping Data

---

### Decomposition

```{r, results = "asis", echo = FALSE}
req_read("Computational Thinking ", url = "http://www.lukefabish.com/computational-thinking/", type = "Required")
```

Please read the **Decomposition** and **Hidden Assumptions** sections only - stop when you reach the beginning of the **Turn it into a recipe (let’s make an algorithm)** section.

### Tidy Data and Reshaping

```{r, results = "asis", echo = FALSE}
req_vid("Reshaping Data (Pivoting)", type = "Required", url = "https://www.youtube.com/embed/4ZIFb1BoIB4")
```

```{r, results = "asis", echo = FALSE}
req_read("R4DS 12.1-12.3: Tidy Data", type = "Recommended", url = "https://r4ds.had.co.nz/tidy-data.html")
```


```{r, results = "asis", echo = FALSE}
req_read("R4DS 12.4-12.5: Separate/Unite and Missing Values ", type = "Optional", url = "https://r4ds.had.co.nz/tidy-data.html")
```

(A few more tricks for data cleaning/wrangling, if you're interested.)



---

```{r, results = "asis", echo = FALSE}
checkin("Pivoting")
```


**Question 1:** Create a new dataset called `cereals_3`, that has three columns:

* The name of the cereal

* A column called "Nutrient" with values `protein`, `fat`, or `fiber`.

* A column called "Amount" with the corresponding amount of the nutrient.

**Question 2:** Why didn't we have to add a `rowid` to pivot wider in this case?

```{r, results = "asis", echo = FALSE}
checkin("Decomposition")
```

Cereals in this dataset are placed on shelf 1, 2, or 3.  We would like to know if these cereal placements correspond to different nutritional values; for example, perhaps sugary cereals made for children are on a lower shelf.

Create a new dataset called `cereals_4`, that has four columns:

* The name of the manufacturer

* The mean amount of sugar in cereals on shelf 1.

* The mean amount of sugar in cereals on shelf 2.

* The mean amount of sugar in cereals on shelf 3.

---

## Joining data

---

```{r, results = "asis", echo = FALSE}
req_read("Mutating Joins", url = "https://www.garrickadenbuie.com/project/tidyexplain/", type = "Required")
```


```{r, results = "asis", echo = FALSE}
req_read("R4DS Chapter 13: Relational Data", url = "https://r4ds.had.co.nz/relational-data.html", type = "Recommended")
```

---

```{r, results = "asis", echo = FALSE}
checkin("Practice with Joins")
```


The following code creates three datasets:

```{r}
data1 <- data.frame(
  professor = c("Bodwin", "Glanz", "Carlton", "Sun"),
  undergrad_school = c("Harvard", "Cal Poly", "Cal Poly", "Harvard"),
  grad_school = c("UNC", "Boston University", "UCLA", "Stanford")
)

data2 <- data.frame(
  professor = c("Bodwin", "Glanz", "Carlton"),
  Stat_331 = c(TRUE, TRUE, TRUE),
  Stat_330 = c(FALSE, TRUE, TRUE),
  Stat_431 = c(TRUE, TRUE, FALSE)
)

data3 <- data.frame(
  course = c("Stat_331", "Stat_330", "Stat_431"),
  num_sections = c(8, 3, 1)
)

```

Here is what they look like once created:

```{r, echo = FALSE}
data1

data2

data3
```

These datasets contain information about four Cal Poly professors, their educational history, the classes they are able to teach, and the number of sections of each class that need to be assigned.

We'd like to use all three datasets to figure out the best assignment of courses.

**Question 1: Warm-up** 

a) Combine datasets 1 and 2 to make this:

```{r, echo = FALSE}
data1 %>% full_join(data2)


```

b) Combine datasets 1 and 2 to make this:

```{r, echo = FALSE}
data1 %>% right_join(data2)


```

c) Combine datasets 2 and 3 to make this:

```{r, echo = FALSE}
data2 %>%
  pivot_longer(Stat_331:Stat_431, 
               names_to = "course",
               values_to = "can_teach") %>%
  full_join(data3)
```


**Question 2: Solve the puzzle**

We would like to assign teachers for next year, with the following guidelines:

1. Each professor should be given 4 sections.

2. At least half of all 331 sections should be taught by Cal Poly graduates.

3. More than half of all 330 sections should be taught by UCLA graduates.

4. Each professor should teach at least 2 different classes.

Combine the three datasets into a single data frame that helps you find
the answer to the question:

*How many sections of Stat 331 will Dr. Glanz teach?*

