---
title: "Git, GitHub, and R"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = here::here("Images", "icons")) 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(couRsework)
library(flair)
library(lubridate)
```


```{css, echo = FALSE}
img {
  vertical-align:middle;
}

body {
  text-shadow:none;
}

```



**GitHub** is a site for sharing and collaborating on code, and for careful tracking
of changes as you progress.  As of right now, it is the best way to organize your
work in **R**, to showcase it to  employers, or to share it with collaborators.

Using **Git** and **GitHub** takes some getting used to, and can be frustrating
at times even to experts.  We will use it only in the most basic form in this course.




```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "45 min", 
                   reading_time = "10 min", 
                   work_time = "30 min", 
                   num_activities = 4)
```

```{r, results = "asis", echo = FALSE}
extra_rec("Git and GitHub")
```

* [This Video](https://www.youtube.com/watch?w3jLJU7DT5E) provides a cheesy but 
cute introduction to GitHub.
* [GitHub Desktop](https://desktop.github.com/) is a nice client for interfacing 
with GitHub.  I highly recommend downloading it if you have not used GitHub before.
* [This Video](https://www.youtube.com/watch?v=KjLycV1IWqc) summarizes the basics
of GitHub with R, and is a good reference if you return to this material after
your first pass through.

---

## Introduction to Git and GitHub

```{r, results = "asis", echo = FALSE}
req_vid("Intro to Git and Github", type = "Required", url = "https://www.youtube.com/embed/BCQHnlnPusY")
```


```{r, results = "asis", echo = FALSE}
checkin("Make a GitHub Account")
```

Create an account on [GitHub.com](GitHub.com), if you have not already.

Ideally, your username will be related to your real name, e.g. "kbodwin" rather 
than "SuperKrazy2000".  However, this is not required; choose whatever username 
you prefer.  


```{r, results = "asis", echo = FALSE}
checkin("Make a repository")
```

Create a **private** repository called "Test Repo".

Create a new text file called "My favorite".  Copy-paste your favorite quote or
poem into the file.

Take a screenshot of your repository.  It should look similar to 9:43 of the video.
 
---

## Branches and Pull Requests

```{r, results = "asis", echo = FALSE}
req_vid("Branching", type = "Required", url = "https://www.youtube.com/embed/oPpnCh7InLY")
```


```{r, results = "asis", echo = FALSE}
checkin("Make a branch and pull request")
```

Create a new branch of your "Test Repo" repository.  
Name this branch after your favorite animal.

In this branch, replace a word in your text file with the animal's name.

Create a **Pull Request** to your `master` branch from your new branch.

Take a screen shot of your completed pull request.  It should look like 8:14 of the video.

---

## Git and RStudio

```{r, results = "asis", echo = FALSE}
req_vid("GitHub and RStudio", type = "Required", url = "https://www.youtube.com/embed/qcjpHFwCugE")
```

```{r, results = "asis", echo = FALSE}
checkin("Access GitHub from RStudio")
```

Follow the instructions in the video to make a new R Project that is also a 
GitHub Repository.

Take a screenshot showing yourself committing to a new branch. It should look
like 8:26 of the video.

**Note:** There are a lot of fiddly computer details that might prevent this from
going as smoothly as in the video.  This doesn't necessarily mean you are doing
something wrong!  If you hit a snag, please contact Dr. B. to get help.

---

## Renaming the `master` branch

```{r, results = "asis", echo = FALSE}
req_read("Renaming the master branch", type = "Required", url = "http://www.kapwing.com/blog/how-to-rename-your-master-branch-to-main-in-git/")
```

There is no Check-In for this reading; I ask only that you read the blog post and
think carefully about the issues.  Please do not skip this section.

If you do feel that you wish to rename your master branch, I am happy to help
you through the process.
