---
title: "Package Basics"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = here::here("Images", "icons")) 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(couRsework)
library(flair)
library(lubridate)
```


```{css, echo = FALSE}
img {
  vertical-align:middle;
}

body {
  text-shadow:none;
}

```


In this short lesson, we will learn about the structure of R Packages.

You will never be asked to write a full package in this class, but in the next
three weeks, you will sometimes submit your work by writing functions in a package 
that I give to you.  Thus, we need to understand a little bit about how packages work.

```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "0 min", 
                   reading_time = "10 min", 
                   work_time = "15 min", 
                   num_activities = 1)
```

```{r, results = "asis", echo = FALSE}
extra_rec("")
```

* [R Packages](http://r-pkgs.had.co.nz/intro.html) by Hadley Wickham - more detailed
and complex than you need, but a great resource if you ever go further with R packages.

---

## Basics of Packages

In principle, an R Package is nothing more than a folder with a very specific structure, 
which allows R to recognize it as a "Package".  Minimally, to be a package, a folder
must contain:

1. A subfolder named `/R/`, containing `.R` code files.
2. A text file named `DESCRIPTION`, which contains information about the package 
in a *very* specifically formatted way.
3. A text file named `NAMESPACE`, which contains a list of the functions that 
the package makes available.

(If you're thinking to yourself, *"Boy, that sounds like a lot of annoyingly specific
formatting!"* - you're not wrong.  Fortunately, there are automatic ways to make 
these files, if you ever write your own packages.)

Any folder with that structure can, in principle, be **installed** into R as a
package.  Typically, you will be installing packages from one of three sources:

#### 1. `install.packages("something")`

If you are able to install a package directly using `install.packages()`, that
means the package has been accepted to the [Comprehensive R Archive Network](https://cran.r-project.org/),
or **CRAN**.  

The minimal folder structure is not sufficient to be accepted to CRAN; these 
packages must meet many strict guidelines about documenting and testing the package's functions.

#### 2. `install_github("username/something")`

You may have come across packages you wanted to use that could only be installed
using the `install_github` function (from the `remotes` or `devtools` package).

In principle, anyone can put a package-structured folder on GitHub, and if the repository
is public, anyone else can install that package.

In practice, it's good to have your package meet CRAN-like levels of careful 
documentation before you share it with the public.


#### 3. Installing from your personal sources

If you use R heavily in the future, you may find it useful to write packages
just for yourself.  For example, if you find yourself using the same small "helper"
functions over and over, it can be nice to simply load a package rather than re-run
them for every project.

We won't worry about installing from sources for this class.

---


```{r, results = "asis", echo = FALSE}
checkin("Basic Package Structure")
```


Visit [this GitHub repository](https://github.com/Cal-Poly-Advanced-R/meangirls), 
where our good friend Regina George has made a package.

Answer the following questions about Regina's package by clicking around in her files.

1. What is the name of this package?
2. Besides Regina, who is listed as an author of this package?
3. Which other packages does Regina's package depend on?  (Hint: Which packages 
does this one *import*?)
4. Which functions are defined in the file `insults.R`?
5. In the documentation comments above the functions, what does `@param` indicate?
6. In the documentation comments above the functions, what does `@importFrom` do?
