---
title: "Final Exam"
author: 'Stat 331: Summer 2020'
output: rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE)
```

```{r, eval = TRUE, include = FALSE}
library(tidyverse)
library(infer)
```

```{r}
np_trails <- read_csv(here::here("11_Final_Exam", "national_park_trails.csv"))
```

## Introduction

This exam is due at **8:00am** on **Friday, August 28th**.
Late exams will be docked by a penalty of at least 50%, possibly more.
**Please do not turn your exam in late!!!**

Please create a new R Markdown file for each of the three parts of this exam.
Make sure your files are set up to display all your source code when knitted.

Clearly mark the questions, so that I can easily find your answers to grade them.
Please do NOT use nonstandard R Markdown templates for this exam.

Although you are not directly graded on data exploration, cleaning, and adjusting, 
you will certainly need to do this to properly address the questions.

### Policies

You may use any resource at all for this exam, with the exception of communicating
with other human beings, in or out of the class.

If you copy substantial code snippets (say, more than three lines of code) from sources on 
the internet, that is okay, but please make sure you cite your source.

### The Data

The dataset for this exam concerns hiking trails in America's National Parks system.
Data is taken from the *AllTrails* app, which allows users to input new trails
and rate existing ones.
You can find documentation on the dataset from its creator [here](https://www.kaggle.com/planejane/national-park-trails).

Variables in the data are:

* `trail_id`: An ID number for the trail
* `name`: Name of the trail
* `area_name`: National Park where the trial is located
* `city_name`: City where the trail is located
* `state_name`: State where the trail is located
* `country_name`: Country where the trail is located
* `_geoloc`: Latitude and longitude of trail
* `popularity`: A score from 0 to 100 of how often the trail is used.
* `length`: Distance covered by trail (in meters).
* `elevation_gain`: How much uphill is on the trail (in meters).
* `difficulty_rating`: Average user rating of trail difficulty
* `route_type`: Type of trail (loop, out and back, or point to point)
* `visitor_usage`: (Honestly, I'm not sure what this one is.)
* `avg_rating`: The average visitor rating of the trail (1 to 5 stars)
* `num_reviews`: The total number of reviews on the app for that trail
* `features`: Notable elements about the trail
* `activities`: Common uses for the trail
* `units`: **This variable is a lie!** All the lengths are in meters.




## Part One: Short answer questions

Answer each of the following using code.  The output of the code should clearly 
and directly answer the question.  You do not need to provide any text answers;
only code and output.  For example, if the question is, "What is the longest bill
length of any penguin in the dataset?", you might provide:

```{r, echo = TRUE, eval = TRUE}
library(palmerpenguins)

penguins %>% 
  top_n(1, bill_length_mm) %>%
  select(bill_length_mm)
```

#### Question 1.1  [10 points]

Which National Park contains the most trails with a 5-star average rating?

```{r}
np_trails %>%
  filter(avg_rating == 5) %>%
  count(area_name) %>%
  top_n(1, n)
```


#### Question 1.2 [10 points]

What are the three longest trails in this dataset, in miles?

```{r}
np_trails %>%
  mutate(
    length_miles = length*3.3/5280
  ) %>%
  top_n(3, length_miles) %>%
  select(name, area_name, length_miles)
```

#### Question 1.3 [10 points]

What is the most popular trail in each National Park?

```{r}
np_trails %>%
  group_by(area_name) %>%
  summarize(
    best_trail = name[popularity = max(popularity)])
```



#### Question 1.4 [10 points]

How many National Parks contain the word "Canyon" somewhere in their name?

```{r}
np_trails %>%
  distinct(area_name) %>%
  pull(area_name) %>%
  str_detect("[Cc]anyon") %>%
  sum()
```

#### Question 1.5 [20 points]

What are the five common words used in trail names?

Do not include besides words that mean "trail", like "trail", "route", "path", "loop".

Do not include connector words like "to", "and", "with", etc.

```{r}
trail_words <- np_trails %>%
  pull(name) %>%
  str_extract_all("\\w+") %>%
  unlist() 

word_counts <- tibble(
  word = trail_words
) %>%
  count(word) %>%
  arrange(desc(n))
```

#### Question 1.6 [20 points]

Write a function that takes as input the dataset and the name of a National Park, 
and the following optional arguments:

* A minimum trail length (in miles)
* A maximum trail length (in miles)
* A minimum trail rating
* The type of trail

The function should return all the trails that fit the supplied criteria. You **may**
assume that the variable names of the inputted dataset match the `np_trails` dataset.

Run the following code to show your function works:

```{r}
recommend_trails <- function(data, park_name, 
                            min_len = 0, max_len = Inf, min_rating = 0, type = NULL) {
  
  if (!is.null(type)) {
    data <- data %>%
      filter(route_type == type)
  }
  
  data %>%
    mutate(
      length = length*3.3/5280
    ) %>%
    filter(
      area_name == park_name,
      length > min_len,
      length < max_len,
      avg_rating > min_rating
    )
  
  
}
```


```{r, echo = TRUE}
np_trails %>%
  recommend_trails("Haleakala National Park", min_len = 10, min_rating = 4)
```


#### Question 1.7 [20 points]

Use your function from Question 1.6, along with a `map` or `apply` function, to 
recommend to me all the trails that are:

* 10-20  miles long  (good for an overnight backpacking trip!)

* Rated at least a 4.5

* Loop trails (I don't like to retrace my steps)

* In one of the parks I can drive to from San Luis Obispo:  Sequoia National Park, 
Joshua Tree National Park, Lassen Volcanic National Park, Kings Canyon National Park, 
Pinnacles National Park, Death Valley National Park, Redwood National Park

```{r}
parks <- c("Sequoia National Park", "Joshua Tree National Park", "Lassen Volcanic National Park", "Kings Canyon National Park", "Death Valley National Park", "Redwood National Park")

map_dfr(parks, ~recommend_trails(np_trails, .x, 10, 20, 4.5, "loop"))
```



## Part Two: Comparing Parks

The two National Parks closest to San Luis Obispo are **Kings Canyon** and **Sequoia**.
These two parks border each other, and are managed under a single forest service;
sometimes, they are simply referred to together as "SeKi".

But are they actually meaningfully different parks?  Let's find out!

In this section, please show your code and output, but also include appropriate
text discussion to interpret the output and answer the question.

#### Question 2.1 [50 points]

Is there evidence that one of these two parks has better trails?

That is:

a) Is there a difference in **mean** trail rating in the two parks? Answer using
a *parametric* hypothesis test.

b) Is there a difference in **median** trail rating in the two parks?  Answer by
creating a bootstrapped confidence interval for each park.

```{r}
seki <- np_trails %>%
  filter(str_detect(area_name, "Kings Canyon|Sequoia")) 

seki %>%
  t_test(elevation_gain ~ area_name)
```

```{r}
seki_boots <- modelr::bootstrap(seki, n=500)

seki_boots <- seki_boots %>%
  mutate(
    se_meds = map_dbl(strap, ~ data.frame(.x) %>% filter(str_detect(area_name, "Sequoia")) %$% median(avg_rating)),
    kc_meds = map_dbl(strap, ~ data.frame(.x) %>% filter(str_detect(area_name, "Kings")) %$% median(avg_rating))
  )

summary(seki_boots)
```



#### Question 2.2 [50 points]

Do these parks have different steepness of trails?

Create a linear model, and a corresponding plot, that predicts the elevation 
gain of a trail to its length.

Interpret the results to answer:

* What is the average steepness (gain/length) of a trail in Kings Canyon?

* What is the average steepness of a trail in Sequoia?

* Is there evidence that these steepnesses are significantly different?

```{r}
my_mod <- lm(elevation_gain ~ length*area_name, data = seki)


seki %>%
  ggplot(aes(x = length, y = elevation_gain, color = area_name)) + 
  geom_point() +
  stat_smooth(method = "lm")
```


## Part Three: Open-Ended [100 points]

Create a short writeup (~1 page, including figures) to address the following
question:

**What features or activities make a park or trail well-liked by hikers?**

You may interpret this question however you like:  feel free to study something
very specific related to the question, such focusing on a single park or a single
feature/activity.

Your writeup must include at least one visualization and at least one statistical
analysis.

