---
title: "Functional Programming"
author: ""

output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      
---
class: center, middle, invert
# Iteration


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(flair)
library(palmerpenguins)
library(glue)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo(
  primary_color = "#26116c",
  secondary_color = "#fd5e53",
  title_slide_text_color = "#fff8e7",
  header_font_google = google_font("Roboto"),
  text_font_google   = google_font("Roboto Condensed"),
  code_font_google   = google_font("Droid Mono")
)
```
---

# Iteration

- What if you want to do something a whole bunch of times, or to every item in a list?  

```{r}
cat("99 bottles of pop on the wall, 99 bottles of pop.  
      Take one down, pass it around, 98 bottles of pop on the wall.")

cat("98 bottles of pop on the wall, 98 bottles of pop.  
    Take one down, pass it around, 97 bottles of pop on the wall.")

###.... and so on...
```

---
# We could write a function...

```{r}
sing_about_pop <- function(num) {
  
  
  paste(num,
    "bottles of pop on the wall,",
    num,
    "bottles of pop.",
    "Take one down, pass it around,",
    num - 1,
    "bottles of pop on the wall.")
  
}
```

---
# We could write a function...

```{r}

sing_about_pop <- function(num) {
  
  
  glue("{num} bottles of pop on the wall, {num} bottles of pop.
       Take one down, pass it around, {num - 1} bottles of pop on the wall.")
  
}
```

---
# This helped!

```{r}
sing_about_pop(99)
sing_about_pop(98)
```


---
# But it is still repetitive...

```{r}
pop_song <- paste(
  sing_about_pop(99),
  sing_about_pop(98),
  sing_about_pop(97),
  sing_about_pop(99),
  sing_about_pop(98)
)

pop_song %>% cat(sep = "\n")
```


---
# Old School: For loops

```{r}
pop_song <- c()

for (i in 99:95) {
  
  pop_song <- c(pop_song, sing_about_pop(i))

}

pop_song %>% cat(sep = "\n")
```

---
# Old School: For loops

**Pro:**

* Classic programming approach; easy to understand

* Can change values at each loop step.

--


**Con:**

* A little messy to read the code

* Don't fit "nicely" inside other functions or pipelines.


---
# Better: The `apply` functions

```{r}

sapply(99:95, sing_about_pop) %>%
  cat(sep = "\n")

```
---
# Better: The `apply` functions

- Pay attention to what type of input they *take* and what type of output they *give*

- `sapply` takes a vector, returns a vector

- `lapply` takes a list, returns a list

- `apply` takes a dataset and a dimension (1 = row, 2 = col), returns a vector or list

---
# Best:  The `map` functions from `purrr`


```{r}

map_chr(99:95, sing_about_pop) %>%
  cat(sep = "\n")

```

---
# Best:  The `map` functions from `purrr`

- `map` always returns a **list**

- `map_chr` always returns a **character vector**  (or an error)

- What do you think `map_dbl` returns?

---
class: inverse, center, middle

# Mapping with multiple arguments

---

# What if I want some options in my function?

```{r}
sing_about_drinks <- function(num, drink, container = "bottles") {
  
  
    
  glue("{num} {container} of {drink} on the wall, {num} {container} of {drink}.
       Take one down, pass it around, {num - 1} {container} of {drink} on the wall.")
  
}
```

---
# `sapply` has trouble!

```{r, error = TRUE}

sapply(99:95, sing_about_drinks(drink = "beer"))

```
---

# You can kind of get around this...

(This is called a **headless function**)

```{r app_2, include = FALSE}

sapply(99:95, function(x) sing_about_drinks(x, drink = "beer")) %>%
  cat(sep = "\n")

```

```{r, echo = FALSE}
decorate("app_2") %>%
  flair("function(x)") %>%
  flair("x")
```

---
# You can also do this in `map` functions

```{r map_1, include = FALSE}

map_chr(99:95, function(x) sing_about_drinks(x, drink = "beer")) %>%
  cat(sep = "\n")

```

```{r, echo = FALSE}
decorate("map_1") %>%
  flair("function(x)") %>%
  flair("x")
```


---

# But `map` has a cool shortcut!

- Note the "~" in front of the function - this is important!

- Note the "." before the "x" - this is important!

```{r map_2, include = FALSE}

map_chr(99:95, ~sing_about_drinks(.x, "beer")) %>%
  cat(sep = "\n")

```

```{r, echo = FALSE}
decorate("map_2") %>%
  flair("~") %>%
  flair(".x")
```


---
class: center, middle, inverse

# But what if I want to iterate over multiple options?

---
# Multiple iteration

```{r, echo = FALSE}

data <- data.frame(num = 99:95, drink = c("pop", "beer", "water", "lemonade","milk"))

```

```{r}
head(data)
```

---
# Use `map2`!

```{r map2}
map2_chr(data$num, data$drink, ~sing_about_drinks(.x, .y, "cans")) %>%
  cat(sep = "\n")
```

---
class: center, middle, inverse

# I demand even more options!
---
# Mapping over many options

```{r, echo = FALSE}

data <- data.frame(num = 99:95, 
                   drink = c("pop", "beer", "water", "lemonade","milk"),
                   container = c("bottles", "cans", "glasses", "beyonces", "cartons"))

```

```{r}
head(data)
```

---
# Mapping over many options

`pmap` takes a list or data frame as input.

```{r pmap}
pmap_chr(data, ~sing_about_drinks(..1, ..2, ..3)) %>%
  cat(sep = "\n")
```

---
# Mapping over many options

Since there are only 3 variables, and we want them to go into the arguments of
`sing_about_drinks` in order...

```{r}
pmap_chr(data, sing_about_drinks) %>%
  cat(sep = "\n")
```

---
# Using map with the pipe

The `map` functions work very nicely with `mutate()` and similar pipe-friendly functions.

```{r pens1, include = FALSE}
penguins %>%
  mutate(
    smaller_dimension = min(bill_length_mm, bill_depth_mm)
  ) %>%
  select(bill_length_mm, bill_depth_mm, smaller_dimension)

```

```{r, echo = FALSE}
decorate("pens1") %>%
  flair_rx("min.*\\)")
```


---
# Using map with the pipe

The `map` functions work very nicely with `mutate()` and similar pipe-friendly functions.

```{r pens2, include = FALSE}
penguins %>%
  mutate(
    smaller_dimension = min(bill_length_mm, bill_depth_mm, na.rm = TRUE)
  ) %>%
  select(bill_length_mm, bill_depth_mm, smaller_dimension)

```

```{r, echo = FALSE}
decorate("pens2") %>%
  flair("na.rm = TRUE")
```


---
# Using map with the pipe

The `map` functions work very nicely with `mutate()` and similar pipe-friendly functions.

```{r pens3, include = FALSE}
penguins %>%
  mutate(
    smaller_dimension = map2_dbl(bill_length_mm, bill_depth_mm, min)
  ) %>%
  select(bill_length_mm, bill_depth_mm, smaller_dimension)

```

```{r, echo = FALSE}
decorate("pens3") %>%
  flair("map2_dbl") %>%
  flair("min")
```



---
class: inverse, center, middle

# The best iteration is no iteration

---
# Vectorized functions

## What does it mean to be vectorized?

```{r}
add_odds <- function(x, y){
  
  if(x %% 2 == 1 & y %% 2 == 1){
    
    return(x + y)
    
  } else {
    
   return(0)
    
  }
  
}

```

---

# What does it mean to be vectorized?

```{r}

add_odds(1, 3)

add_odds(2, 4)


```

---

# What does it mean to be vectorized?

```{r, error = TRUE }

first_nums <- c(1, 2)
next_nums <- c(3, 4)

add_odds(first_nums, next_nums)

```

---

# Vectorized Functions

`add_odds` is not a vectorized function!

```{r, eval = FALSE}
add_odds <- function(x, y){
  
  if(x %% 2 == 1 & y %% 2 == 1){
    
    return(x + y)
    
  } else {
    
   return(0)
    
  }
  
}

```

---

# Vectorized Functions

We can avoid the error by iteration, but this is slow!

```{r, error = TRUE }

first_nums <- c(1,2)
next_nums <- c(3,4)

map2_dbl(first_nums, next_nums, add_odds)

```

---
# Vectorized function

A better (but harder!) approach is to make the function **truly** vectorized:

 
```{r}
add_odds <- function(x, y){
  
  is_odd <- x %% 2 == 1 & y %% 2 == 1
  
  answers <- x + y
  
  answers[!is_odd] <- 0
  
  return(answers)
  
}
```

```{r}

add_odds(first_nums, next_nums)

```



---

# Function design: Things to think about

- What does your function take as INPUT?

- What does your function give as OUTPUT?

--

- How can we write it to be friendly with `apply` or `map`?

--

- Do you REALLY need to use `apply` or `map`, or is it vectorized?

--

- Bonus: How can we write it to be friendly with `%>%`?





