---
title: "Objects in R"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = "../../Images/icons") 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(couRsework)
library(flair)
```

In this lesson, you'll learn some formal definitions and vocabulary for objects in R.

When you are finished, you should be able to...

* Identify the **structure** of an object

* Identify the **data type(s)** of an object.

* Understand the *required* and *optional* arguments in a function.

* Troubleshoot common error messages


```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "40 min", 
                   reading_time = "15 min", 
                   work_time = "5 min", 
                   num_activities = 2)
```

---

## Types and Structures

```{r, results = "asis", echo = FALSE}
req_vid("Intro to Object Types and Structures", type = "Required", url = "https://www.youtube.com/embed/abHbBykMWdI")
```

```{r, results = "asis", echo = FALSE}
link_slides("Slides/Types_Structures.html")
```

```{r, results = "asis", echo = FALSE}
req_read("More Detail on Object Types and Structures", type = "Optional", url = "https://ramnathv.github.io/pycon2014-r/learn/structures.html")
```


---


```{r, results = "asis", echo = FALSE}
checkin("Data Structures")
```

This lecture did **not** mention the important structure of **data frames**.

That's because, in essence, a **data frame** is simply a special **list** - with a few extra restrictions on the list format.  

Think about the datasets you have already worked with. Which of the following restrictions on lists do you think are needed for the list to be a **data frame**?

a) The elements of the list must all be vectors of the same length.
b) The elements of the list must all be the same data type.
c) The elements of the list must all have no missing values.
d) The elements of the list must all have names.

---

## Functions

```{r, results = "asis", echo = FALSE}
req_vid("Intro to Functions", type = "Required", url = "https://www.youtube.com/embed/Cp_9y39ZRRM")
```


```{r, results = "asis", echo = FALSE}
link_slides("Slides/Functions.html")
```


```{r, results = "asis", echo = FALSE}
req_read("Basics of R", type = "Required", url = "https://r4ds.had.co.nz/workflow-basics.html")
```

---


```{r, results = "asis", echo = FALSE}
checkin("Troubleshooting")
```

What is the reason for the error below?  How would you fix it?

```{r, error = TRUE}
matrix(c("a", "b", "c", "d"), num_row = 2)
```
