---
title: "Distributions and Simulation"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = here::here("Images", "icons")) 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(couRsework)
library(flair)
```



```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "30-60 min", 
                   reading_time = "10 min", 
                   work_time = "45 min", 
                   num_activities = 3)
```

---

## Review: Statistical Distributions


```{r, results = "asis", echo = FALSE}
req_vid("Review of Statistical Distributions", type = "Required", url = "https://www.youtube.com/embed/KddbZr-ssZc")
```

```{r, results = "asis", echo = FALSE}
req_tutorial("Interactive Random Variable Creator", type = "Optional", url = "https://seeing-theory.brown.edu/probability-distributions/index.html#section1")
```

---

## Simulating in R

```{r, results = "asis", echo = FALSE}
req_vid("Simulating Data in R", type = "Required", url = "https://www.youtube.com/embed/LNg7kQbXvU0")
```

```{r, results = "asis", echo = FALSE}
req_read("RProg Ch 20: Simulation", type = "Optional", url = "https://bookdown.org/rdpeng/rprogdatascience/simulation.html")
```

---

```{r, results = "asis", echo = FALSE}
checkin("The r, p, q, and d functions")
```

Try to predict what the following outputs will be **WITHOUT** running the code
in R.  Drawing pictures of the relevant distributions may help.

(Yes, it is very easy to "cheat" on this question. But this is for your practice,
and I recommend you give it some thought.)

```{r, eval = FALSE}

## a
punif(0.674)

## b

pnorm(2)
qnorm(.975)

## c

pchisq(0, df = 12)

## d

qt(10, df = 16)

## e

dbinom(2, size = 2, prob = .4)
pbinom(1, size = 2, prob = .6)


```

---
## Setting the Seed

Functions like `rnorm` and `runif` rely on something called **pseudorandomness**.
Because computers can never be *truly* random, complicated processes are implemented
to make "random" number generation be so unpredictable as to behave like true
randomness.

This means that projects involving simulation are harder to make **reproducible**.
For example, here are two identical lines of code that give different results!

```{r}
runif(1)
```


```{r}
runif(1)
```


Fortunately, pseudorandomness depends on a **seed**, which is an arbitrary number
where the randomizing process starts.  Normally, R will choose the seed for you,
from a pre-generated vector:

```{r}
head(.Random.seed)
```


However, you can also choose your own seed using the `set.seed()` function. This
guarantees your results will be consistent:

```{r}
set.seed(15)
runif(1)
```


```{r}
set.seed(15)
runif(1)
```

(Of course, it doesn't mean the results will be the same in every subsequent run, 
unless you reset the seed in between each line of code!)

```{r}
set.seed(15)
runif(1)
runif(1)
```


It is very important to always set a seed at the **beginning** of an R Markdown document
that contains any random steps, so that your knitted results are consistent.

Note, though, that this only guarantees your knitted results will be the same
*if the code has not changed*.

Changing up any part of the code will re-randomize everything that comes after it:

```{r}
set.seed(15)
runif(1)
runif(1)
```

```{r}
set.seed(15)
rnorm(1)
runif(1)
```

---

## Preparing for the Lab

---

### The Central Limit Theorem

If you do not recall the Central Limit Theorem from you previous classes - or if
you would benefit from a refresher:


```{r, results = "asis", echo = FALSE}
req_vid("The CLT", type = "Optional", url = "https://www.youtube.com/embed/YAlJCEDH2uY")
```

---

```{r, results = "asis", echo = FALSE}
checkin("The CLT")
```

Explain the idea behind the CLT in your own words, using as little jargon/vocabulary
as you can.

---

### Plotting distributions

Here is the code that made one of the plots from the lecture video:

```{r}
my_samples <- data.frame(x = rchisq(1000, df = 5))

ggplot(my_samples, aes(x)) + 
  geom_histogram(bins = 40, aes(y = ..density..), fill = "grey", col = "white") + 
  stat_function(fun= ~dchisq(.x, df = 5), col = "cornflowerblue", lwd = 2) +
  theme_classic()
```

---

```{r, results = "asis", echo = FALSE}
checkin("Plotting")
```

Re-create this plot from the lecture slides:
(The two colors are "cornflowerblue" and "deeppink")

```{r, echo = FALSE}
my_samples <- data.frame(x = rt(1000, df = 5))

ggplot(my_samples, aes(x)) + 
  geom_histogram(bins = 40, aes(y = ..density..), fill = "grey", col = "white") + 
  stat_function(fun= function(x) dt(x, df = 5), col = "cornflowerblue", lwd = 2) +
  stat_function(fun = dnorm, col = "deeppink", lwd = 1) +
  theme_classic()
```

