---
title: "Reproducibility: Directories, Data, and Deliverables"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = "../../Images/icons") 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(couRsework)
library(flair)
```

The theme of this lesson is good management of your files and data.

You will learn how to:

* Identify **folders** and **paths**.

* Create and use **R Projects**.

* **Load data** into R from files.

* Deal with package challenges.

* Create R Markdown documents with good **reproducible principles**.


```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "40 min", 
                   reading_time = "10-20 min", 
                   work_time = "10 min", 
                   num_activities = 2)
```

---

## Directories, etc.

```{r, results = "asis", echo = FALSE}
req_vid("Directories, Paths, Projects", type = "Required", url = "https://www.youtube.com/embed/fKBXX3C_WJU")
```

```{r, results = "asis", echo = FALSE}
req_read("Workflow and Projects", type = "Optional", url = "https://r4ds.had.co.nz/workflow-projects.html")
```

---

```{r, results = "asis", echo = FALSE}
checkin("R Projects")
```

Take a screenshot of your class directory, showing the **R Project** and the folder organization, including a **Data** folder.

---

## Loading Data Into R

### Where does my data live?

```{r, results = "asis", echo = FALSE}

req_vid("Data Locations and the {here} package", type = "Required", url = "https://www.youtube.com/embed/kF_fmHEgUOM")

```

### How do I load the data?

```{results = "asis", echo = FALSE}
req_vid("Loading Data", type = "Required", url = "https://www.youtube.com/embed/WOmXNZm1Jr4")

req_read("Data Import", type = "Recommended", url = "https://r4ds.had.co.nz/data-import.html")

```


---

```{r, results = "asis", echo = FALSE}
checkin("Loading Data")
```

**Question 1**

What line of code did you use - including the `here` function - to read in `ages.csv`?

**Question 2**

What does the `stringsAsFactors` argument do?

**Question 3**

What code did you use to read in `Name` as a factor and `Age` as a character?

**Bonus Question**

Just for fun - did you [recognize the men in this dataset?](https://www.hockey-reference.com/teams/SJS/2020.html)

---

## Packages

```{r, results = "asis", echo = FALSE}
req_vid("Details of Packages", type = "Required", url = "https://www.youtube.com/embed/Ar9leuWF5ws")
```

---

## Reproducible Deliverables

```{r, results = "asis", echo = FALSE}
req_vid("Reproducible Deliverables in R Markdown", type = "Required", url = "https://www.youtube.com/embed/i3MHIBNaZzE")
```

